---
title       : Exponential Distribution
subtitle    : Mean and Variance of Exponential Distribution
author      : Anna Teo
job         : 
framework   : io2012            # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js      # {highlight.js, prettify, highlight}
hitheme     : tomorrow          # 
widgets     : [mathjax]         # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction...

This is a simple application to simulate the exponential distribution in R using rexp(n, lambda) over t times. 


Users have the options to select the parameters:

1. n - number of exponential numbers
2. $\lambda$ (lambda)  - rate parameter
3. t - number of simulations


The application will then display the following:
* the theoretical mean and variance; and 
* plot the distribution of mean and variance

---  

## Exponential Distribution

Theoretical Mean:

* The mean or expected value of an exponentially distributed random variable X with rate parameter lambda is given by

$$E[X] = \frac{1}{\lambda}$$


Theoretical Variance:
* The variance of X is given by

$$Var[X] = \frac{1}{\lambda^2}$$


---

## Demonstration

With the input of n = 10, lambda = 0.2, t = 1000, the application will display:

```{r library, echo=FALSE}
library(ggplot2)
```

```{r simulation_function, echo=FALSE}
simulation <- function(number, lambda, times) {
        mean.s <- NULL
        var.s <- NULL
        for (i in 1:times) {
                value.s <- rexp(number, lambda)
                mean.s <- c(mean.s, mean(value.s)) 
                var.s <- c(var.s, var(value.s))
                }
        df <- data.frame(mean.s, var.s)
        df
        }
```

```{r mutliplot_function, echo=FALSE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r simulation, echo=FALSE, results='hide'}
number <- 10
lambda <- 0.2
times <- 1000
result <- simulation(number, lambda, times)
```
 
Theoretical Mean: E[X] = `r 1/lambda`, Theoretical Variance: Var[X] = `r 1/lambda^2`

```{r plot.mean, fig.height=5, fig.width=12, echo=FALSE}
mean.g <- ggplot(result, aes(x=mean.s)) +
        geom_histogram(alpha = .20, binwidth=0.2, color = "black", fill = "blue", aes(y = ..density..)) +
        geom_vline(aes(xintercept=mean(mean.s, na.rm=T)),color="red", linetype="dashed", size=1) + 
        labs(x = "X") + 
        labs(title = "Distribution of Mean of n exponential numbers") + 
        theme(text = element_text(size=9))

var.g <- ggplot(result, aes(x=var.s)) +
                geom_histogram(alpha = .20, binwidth=2, colour = "black", fill = "green", aes(y = ..density..)) +
                geom_vline(aes(xintercept=mean(var.s, na.rm=T)),color="red", linetype="dashed", size=1) +
                labs(x = "X") + 
                labs(title = "Distribution of Variance of n exponential numbers") + 
                theme(text = element_text(size=9))                        
multiplot(mean.g, var.g, cols = 2)
```

---

## References

[Application](http://lsteo.shinyapps.io/exponential) - hosted on shinyapps.io

[Exponential Distribution](http://en.wikipedia.org/wiki/Exponential_distribution) on Wikipedia
